---
title: "Updating Dogs - error model"
author: "Christoph Völter"
date: "30/05/2023"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
 rm(list=ls())
library (car)
library(lme4)
library(tidyverse)
library(cowplot)
library(glmmTMB)
library(ggsignif)
library("gghalves")
source("./functions/diagnostic_fcns.r")
source("./functions/glmm_stability.r")
source("./functions/boot_glmm.r")
source("./functions/boot_glmm2.r")

#load(file ="exp2_5s_capuchins_GLMM_error_data_wCI.RData")
```
next steps: repeat error model just with self-ordered condition data and with latency from start of trial as additional predictor variable

```{r}
all.data <- read.csv("data/RM_both_sessions.csv") %>%
    mutate(second_choice_correct = ifelse(first_arm==sec_arm, 0, 1),
      third_choice_correct = ifelse(first_arm==third_arm | sec_arm==third_arm, 0,  1),
      fourth_choice_correct = ifelse(first_arm==fourth_arm | sec_arm==fourth_arm | third_arm==fourth_arm, 0,  1),
      fifth_choice_correct = ifelse(first_arm==fifth_arm | sec_arm==fifth_arm | third_arm==fifth_arm | fourth_arm==fifth_arm, 0,  1),
      sixth_choice_correct = ifelse(first_arm==sixth_arm | sec_arm==sixth_arm | third_arm==sixth_arm | fourth_arm==sixth_arm | fifth_arm==sixth_arm, 0,  1),
      seventh_choice_correct = ifelse(first_arm==sevth_arm | sec_arm==sevth_arm | third_arm==sevth_arm | fourth_arm==sevth_arm | fifth_arm==sevth_arm | sixth_arm==sevth_arm, 0,  1),
      eight_choice_correct = ifelse(first_arm==eigth_arm | sec_arm==eigth_arm | third_arm==eigth_arm | fourth_arm==eigth_arm | fifth_arm==eigth_arm | sixth_arm==eigth_arm | sevth_arm==eigth_arm, 0,  1),
  )%>%
  mutate(Number_arms = 8)
```


### Error model
```{r}
number_arms <- 8
str(all.data)
test.data.prep = data.frame(
  all.data[rep(1:nrow(all.data), each = number_arms), c(
    "CDL_name",
    "sex",
    "Age",
    "Number_arms",
    "session",
    "trial_num", 
    "condition"
  )],
  trial.id = rep(1:nrow(all.data), each = number_arms),
  choice.nr.within.trial = rep(1:number_arms, nrow(all.data)),
  choice.where = as.vector(unlist(c(t(
    all.data[, c("first_arm",
                 "sec_arm",
                 "third_arm",
                 "fourth_arm",
                 "fifth_arm",
                 "sixth_arm",
                 "sevth_arm",
                 "eigth_arm")]
  )))),
  choice.correct = as.vector(unlist(c(t(
    cbind(NA, all.data[, c(
      "second_choice_correct",
      "third_choice_correct",
      "fourth_choice_correct",
      "fifth_choice_correct",
      "sixth_choice_correct",
      "seventh_choice_correct",
      "eight_choice_correct"
      )])
  ))))
)
u.trial.id = unique(test.data.prep$trial.id)
last.choice = rep(NA, nrow(test.data.prep))
previous.correct = rep(NA, nrow(test.data.prep))
any.error.earlier = rep(NA, nrow(test.data.prep))
for (i in 1:length(u.trial.id)) {
  sel.data = subset(test.data.prep, trial.id == u.trial.id[i])
  #if(length(unique(sel.data$choice.where[!is.na(sel.data$choice.where)]))<sum(!is.na(sel.data$choice.where)))
  last.choice[test.data.prep$trial.id == u.trial.id[i]][2:number_arms] = unlist(lapply(2:nrow(sel.data), function(x) {
    ires = NA
    if (!is.na(sel.data$choice.where[x])) {
      xx = which(sel.data$choice.where[1:(x - 1)] == sel.data$choice.where[x])
      if (length(xx) > 0) {
        ires = max(xx)
      }
    }
    return(ires)
  }))
  previous.correct[test.data.prep$trial.id == u.trial.id[i]] = c(NA, 1, sel.data$choice.correct[2:(number_arms-1)])
  xx = c(1, sel.data$choice.correct[1:(number_arms-1)])
  xx[is.na(xx)] = 1
  any.error.earlier[test.data.prep$trial.id == u.trial.id[i]] = c("no", "yes")[1 +
                                                                                 cummax(xx == 0)]
}
test.data.prep$last.choice = last.choice
test.data.prep$previous.correct = previous.correct
test.data.prep$any.error.earlier = any.error.earlier
test.data = subset(test.data.prep,
                   choice.nr.within.trial > 1 & !is.na(choice.where))

test.data.prep = subset(test.data.prep,!is.na(choice.where))# 	for(j in 2:nrow(sel.data)){
test.data = c()
choice.id = 0
for (i in 1:length(u.trial.id)) {
  sel.data = subset(test.data.prep, trial.id == u.trial.id[i])
  cup.id = 1:8

  for (j in 2:nrow(sel.data)) {
    choice.id = choice.id + 1
    idata = data.frame(sel.data[j,], cup.id = cup.id, choice.id = choice.id)
    idata$last.choice = unlist(lapply(1:nrow(idata), function(x) {
      xx = which(sel.data$choice.where[1:(j - 1)] == idata$cup.id[x])
      if (length(xx) > 0) {
        return(max(xx))
      } else{
        return(NA)
      }
    }))
    if (j > 2) {
      idata$previous.correct = sel.data$choice.correct[j - 1]
    } else{
      idata$previous.correct = 1
    }
    idata$any.error.earlier = as.numeric(sum(sel.data$choice.correct[1:(j -
                                                                          1)] == 0, na.rm = T) > 0)
    test.data = rbind(test.data, idata)
  }
}

test.data = subset(test.data,!is.na(last.choice))
error.ny = as.numeric(test.data$choice.correct == 0 &
                        test.data$choice.where == test.data$cup.id)
xx = tapply(error.ny, test.data$choice.id, sum)
xx = xx[xx > 0]
lag.from.last.choice = test.data$choice.nr.within.trial - test.data$last.choice
test.data$previous.correct = as.factor(c("no", "yes")[1 + as.numeric(test.data$previous.correct)])
test.data$any.error.earlier = as.factor(test.data$any.error.earlier)
number.cups = as.vector(table(test.data$choice.id)[as.character(test.data$choice.id)])
table(number.cups, test.data$Number_arms)


age_data<-all.data%>%
  #filter(session == 1)%>%
  group_by(CDL_name)%>%
  summarise(Age=mean(as.numeric(Age), na.rm=TRUE))


test.data <-
  data.frame(test.data,
             lag.from.last.choice,
             number.cups,
             # cup.is.edge,
             error.ny)%>%
  select(-Age)%>%
  full_join(age_data)
  

write.csv(test.data, "data/dog_updating_radial_maze_error_data.csv")
```
 choice_where: location (cup id) of current choice
 cup.id: all arms in which an error could be made
 number_cups: number of error possibilities
 lag_from_last_choice: how many choices are in between the current choice and the last choice of this arm

## add duration from start of trial 
```{r}
block_data <- read.csv("data/RM_both_sessions.csv",  na.strings=c("#N/A")) %>% 
  select(Subject = CDL_name, session, block, condition, trial_num)%>%
  mutate(trial_num = as.integer(ifelse(block == 1, trial_num,
                                       ifelse(block == 2, trial_num+3, ""))),
         session = as.numeric(session))%>%
  select(-block)

s1_time_first_err_data <- read.csv("data/First_Session_Time_Analysis_corrected_2.csv",  na.strings=c("#N/A")) %>% 
  mutate(session = 1) %>%
  select(Subject, condition, session, trial_num = trial_number, Trial.start, choice_1 = X1st.choice, choice_2 = X2nd.choice, choice_3 = X3rd.choice, choice_4 = X4th.choice, choice_5 = X5th.choice, choice_6 = X6th.choice, choice_7 = X7th.choice, choice_8 = X8th.choice, choice_9 = X9th.choice)

levels(as.factor(s1_time_first_err_data$Subject))

s2_time_first_err_data <- read.csv("data/Second_Session_Time_Analysis_corrected_2.csv",  na.strings=c("#N/A")) %>% 
  mutate(session = 2) %>%
  select(Subject, session, condition, trial_num = trial_number, choice_1 = X1st.choice, choice_2 = X2nd.choice, choice_3 = X3rd.choice, choice_4 = X4th.choice, choice_5 = X5th.choice, choice_6 = X6th.choice, choice_7 = X7th.choice, choice_8 = X8th.choice, choice_9 = X9th.choice)

levels(as.factor(s2_time_first_err_data$Subject))

time_first_err_data <- s1_time_first_err_data %>%
  full_join(s2_time_first_err_data)%>%
  mutate(choice_2 = choice_2 - choice_1,
         choice_3 = choice_3 - choice_1,
         choice_4 = choice_4 - choice_1,
         choice_5 = choice_5 - choice_1,
         choice_6 = choice_6 - choice_1,
         choice_7 = choice_7 - choice_1,
         choice_8 = choice_8 - choice_1,
         choice_9 = choice_9 - choice_1,
         trial_num = as.integer(ifelse(trial_num>3, trial_num-3, trial_num))
         )%>%
  select(-choice_1, -choice_9, -Trial.start) %>%
  pivot_longer(cols=c(choice_2:choice_8), names_to = "choice", values_to = "duration_from_1st_choice")%>%
  separate(choice, into = c(NA, "choice.nr.within.trial"))%>%
  mutate(choice.nr.within.trial = as.integer(choice.nr.within.trial))%>%
  mutate(Subject = fct_recode(Subject, "Nikita" = "Nikita4"))


missing_time_data<-time_first_err_data%>%
  filter(is.na(duration_from_1st_choice)) %>%
  select(Subject:trial_num)

levels(as.factor(time_first_err_data$trial_num))
```



```{r}
test.data <- read.csv("data/dog_updating_radial_maze_error_data.csv")%>%
  filter(CDL_name!="Ylvie")%>%
  rename(Subject = CDL_name)%>%
  mutate(Age = as.numeric(ifelse(Subject == "Anubis3", 59, ifelse(Subject == "Ace3", 48, ifelse(Subject == "Nikita", 56, Age)))))%>%
  mutate(Subject)%>%
  full_join(time_first_err_data)%>%
  mutate(check_duration = is.na(duration_from_1st_choice))%>%
  anti_join(missing_time_data) #remove trials with NAs for the duration_from_1st_choice variable

table(test.data$Subject, test.data$check_duration)
levels(as.factor(test.data$Subject))
levels(as.factor(test.data$trial_num))
```


```{r}

xx.fe.re = fe.re.tab(
  fe.model = "error.ny~condition + lag.from.last.choice+previous.correct+any.error.earlier+sex+Age+trial_num+duration_from_1st_choice",
  re = "(1|Subject)+(1|choice.id)+(1|trial.id)",
  other.vars = "number.cups",
  data = test.data
)
xx.fe.re$summary

test.data$z.lag.from.last.choice = as.vector(scale(test.data$lag.from.last.choice))
test.data$any.error.earlier = as.factor(test.data$any.error.earlier)
test.data$any.error.earlier.code = as.vector(scale(as.numeric(test.data$any.error.earlier==levels(as.factor(test.data$any.error.earlier))[2]), center=TRUE, scale= FALSE))

test.data$condition <- as.factor(test.data$condition)
test.data$condition.code = as.vector(scale(as.numeric(test.data$condition==levels(as.factor(test.data$condition))[2]), center=TRUE, scale= FALSE))
                                                
test.data$z.trial_number = as.vector(scale(test.data$trial_num))
test.data$z.session = as.vector(scale(test.data$session))
test.data$z.age = as.vector(scale(test.data$Age))
test.data$z.duration_from_1st_choice = as.vector(scale(test.data$duration_from_1st_choice))

```
```{r}
save.image("self_ordered_search_error_analysis.RData")
```


```{r}
mean(test.data$lag.from.last.choice)

mean(test.data$Age)

mean(test.data$trial_num)

sd(test.data$lag.from.last.choice)
sd(test.data$Age)
sd(test.data$trial_num)

levels(as.factor(test.data$Subject))

table(test.data$Subject, test.data$Age)
```

```{r}
contr=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1000000))

test.data$ot=log((1/test.data$number.cups))

full = glmer(
  error.ny ~ condition + z.lag.from.last.choice + z.duration_from_1st_choice + any.error.earlier + sex +
    z.age + z.trial_number + z.session + offset(ot) +
    (
      1 + condition.code + z.lag.from.last.choice + z.duration_from_1st_choice + any.error.earlier.code + z.trial_number + z.session || Subject
    ) +
    (1 | choice.id) + (1 | trial.id),
  family = binomial,
  data = test.data,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1000000))
)

```


```{r}
null = glmer(
  error.ny ~  z.trial_number +  offset(ot) +
    (
      1 + condition.code + z.lag.from.last.choice + z.duration_from_1st_choice + any.error.earlier.code +  z.trial_number + z.session || Subject
    ) +
    (1 | choice.id) + (1 | trial.id),
  family = binomial,
  data = test.data,
  control = contr
)


anova(null, full, test="Chisq")
```

#### Check collinearity
```{r}

 xres=lm(error.ny~condition + z.lag.from.last.choice + any.error.earlier + z.duration_from_1st_choice + sex +
    z.age + z.trial_number + z.session + offset(ot) , data=test.data)

vif(xres)
```



```{r}
drop1_full<-drop1(full, test="Chisq")
drop1_full
```


```{r}
summary(full)
```



```{r}
test.data<-test.data%>%
  mutate(sex.c=as.vector(scale(as.numeric(as.factor(sex), scale=FALSE))))

pred.full.lag=glmer(error.ny~condition.code + lag.from.last.choice + z.duration_from_1st_choice + any.error.earlier.code + sex.c +
    z.age + z.trial_number + z.session + offset(ot) +
    (
      1 + condition.code + z.lag.from.last.choice + z.duration_from_1st_choice + any.error.earlier.code +  z.trial_number + z.session || Subject
    ) +
    (1 | choice.id) + (1 | trial.id),
	family=binomial, data=test.data, control=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1000000)))


```


```{r}
save.image("dog_radial_maze_error_model_workspace.Rdata")
```

```{r}
load("dog_radial_maze_error_model_workspace.Rdata")
boot.ci.full=boot.glmm.pred(model.res=full, excl.warnings=F,
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95)




full.stab=glmm.model.stab(model.res=full, contr=contr, para=T)
table(full.stab$detailed$warnings)
m.stab.plot(full.stab$summary[,-1])


#CIs for plots

pred.full.lag.ci=boot.glmm.pred(model.res=pred.full.lag , resol=100, level=0.95, use="lag.from.last.choice", n.cores="all-1", para=T)

save.image("dog_radial_maze_error_model_workspace_wCI.Rdata")
```

```{r}
error_model_output_table <-
  bind_cols(as.data.frame(summary(full)$coeff),
            drop1_full
           # mm1_choice.ci$ci.estimates
           ) %>%
  select(
    Estimate,
    SE = `Std. Error`,
  #  LowerCI = X2.5.,
 #   UpperCI = X97.5.,
    Chi2 = LRT,
    df = npar,
    p = `Pr(Chi)`,
    z_wald = `z value`,
    p_wald = `Pr(>|z|)`
  ) %>% #
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall = 3))) %>%
  mutate(across(.cols = c(Estimate:Chi2), ~ format(round(.x, 2), nsmall = 2))) %>%
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p = replace(p, p == "0.000", "<0.001"))
error_model_output_table
write.csv(error_model_output_table, file = "saves/error_model_output_table2.csv")
```



```{r}


plot.error.lag.data<-test.data%>%
  mutate(err=error.ny*((number_arms-1)/number.cups))%>%
 # mutate(cup.is.edge=recode(cup.is.edge, "yes"="outer boxes", "no"="inner boxes"))%>%
  group_by(Subject, lag.from.last.choice )%>%
  summarise(mean_err2=mean(err))%>%
  mutate(mean_err = round(mean_err2,2))%>%  
  group_by(lag.from.last.choice )%>%
  add_count(mean_err)

p.lag <-
  ggplot(data = plot.error.lag.data, aes(x = lag.from.last.choice , y = mean_err)) +
  labs(x = "Lag from last choice", y = "Relative proportion of mistakes") +
  #geom_point(size = plot.error.lag.data$n, colour = "black", alpha=0.25) +
  geom_line(
    aes(group = Subject),
    lty = 2,
    alpha = 0.25,
    colour = "black"
  ) +
  geom_boxplot(
    aes(group = lag.from.last.choice),
    alpha = 0.5 ,
    width = 0.4,
    outlier.colour = "transparent",
    coef = 0
  ) +
  geom_count(alpha = 0.25) +
  #   geom_errorbar(aes(x=1,y=pred.full.edge.ci$ci.predicted$fitted[1], ymin=pred.full.edge.ci$ci.predicted$lower.cl[1], ymax=pred.full.edge.ci$ci.predicted$upper.cl[1]), width=.3, col="darkblue") +
  #   geom_point(aes(x=1,y=pred.full.edge.ci$ci.predicted$fitted[1]),  col="darkblue", pch=1) +
  # geom_errorbar(aes(x=2,y=pred.full.edge.ci$ci.predicted$fitted[2], ymin=pred.full.edge.ci$ci.predicted$lower.cl[2], ymax=pred.full.edge.ci$ci.predicted$upper.cl[2]), width=.3, col="darkblue") +
  #   geom_point(aes(x=2,y=pred.full.edge.ci$ci.predicted$fitted[2]), col="darkblue", pch=1) +
  # ylim(0,0.7)+
  theme_bw() +
  theme(
    axis.ticks.x = element_blank(),
    plot.title = element_text(size = 10),
    plot.caption = element_text(size = 10)
  ) +
  scale_x_continuous(breaks = seq(1, 7, 1))


p.lag

ggsave(p.lag, file="graphs/lag_from_last_choice_plot.png", width = 7, height = 5, scale = 0.9)

```






